Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-06-22 article-models-ttdp.tex
revising the rules for evaluation

\begin{align*}
& \inference
    { \jfam{{\Gamma}{A}}{P}
      }
    { \jfam{\Gamma}{\mprd{A}{P}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jfameq{{\Gamma}{A}}{P}{P'}
        }
      { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd{A'}{P'}}
        }
  \\
& \inference
    { \jfam{{{\Gamma}{A}}{P}}{Q}
      }
    { \jfam{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
        }
      { \jfameq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\mprd[\famsym]{A'}{Q'}}
        }
  \\
& \inference
    { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
      }
    { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\slam{A}{Q}{g}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
        }
      { \jtermeq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\slam{A}{Q}{g}}{\slam{A'}{Q}{g'}}
        }
  \\
& \inference
    { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{g}
      }
    { \jterm{{{\Gamma}{A}}{P}}{Q}{\sev{A}{Q}{g}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jtermeq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{g}{g'}
        }
      { \jtermeq{{{\Gamma}{A}}{P}}{Q}{\sev{A}{Q}{g}}{\sev{A'}{Q}{g'}}
        }
\end{align*}

@@

\subsubsection{Dependent products are compatible with the empty context}
The empty context can appear in the domain and in the codomain of the dependent
function type constructor. We have the following inference rules explaining
what happens when the empty context appears in the domain:
\begin{align}
& \inference
    { \jfam{\Gamma}{P}
      }
    { \jfameq{\Gamma}{\mprd{\emptyf}{P}}{P}
      }
  \\
& \inference
    { \jfam{{\Gamma}{P}}{Q}
      }
    { \jfameq{{\Gamma}{P}}{\mprd[\famsym]{\emptyf}{Q}}{Q}
      }
  \\
& \inference
    { \jterm{{\Gamma}{P}}{Q}{g}
      }
    { \jtermeq{{\Gamma}{P}}{Q}{\slam{\emptyf}{Q}{g}}{g}
      }
  \\
& \inference
    { \jterm{{\Gamma}{P}}{Q}{g}
      }
    { \jtermeq{{\Gamma}{P}}{Q}{\sev{\emptyf}{Q}{g}}{g}
      }
\end{align}
We have the following infernece rules explaining what happens when the empty
context appears in the codomain:
\begin{align}
& \inference
    { \jfam{\Gamma}{A}
      }
    { \jfameq{\Gamma}{\mprd{A}{\emptyf}}{\emptyf}
      }
  \\
& \inference
    { \jfam{{\Gamma}{A}}{P}
      }
    { \jfameq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{\emptyf}}{\emptyf}
      }
\end{align}
Now we see that we can compare $\mprd{A}{P}$ with $\mprd[\famsym]{A}{P}$ by
seeing $P$ as a family over $\ctxext{{\Gamma}{A}}{\emptyf}$. We will require
the following inference rule to be valid:
\begin{equation}
\inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd[\famsym]{A}{P}}
    }
\end{equation}
This rule will allow us to omit the annotation indicating the action on families,
which we will do from now on. We also note that because the empty family over
$\ctxext{\Gamma}{A}$ is mapped to the empty family over $\Gamma$ by the
dependent function constructor, we obtain the important special cases of
lambda abstraction and evaluation that
\begin{align*}
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jterm{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}
    }
  \\
& \inference
  { \jterm{\Gamma}{\mprd{A}{P}}{f}
    }
  { \jterm{{\Gamma}{A}}{P}{\sev{A}{P}{f}}
    }
\end{align*}
Thus, we retrieve what is originally meant by lambda abstraction.

@@

\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq
      {\Gamma}
      {\mprd{\ctxext{A}{P}}{Q}}
      {\mprd{A}{\mprd{P}{Q}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}
      {\mprd{\ctxext{A}{P}}{R}}
      {\mprd{A}{\mprd{P}{R}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
    }
  { \jtermeq
      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}
      {\mprd{\ctxext{A}{P}}{R}}
      {\slam{\ctxext{A}{P}}{R}{h}}
      {\slam{A}{\mprd{P}{R}}{\slam{P}{R}{h}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jterm{{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}{\mprd{\ctxext{A}{P}}{R}}{h}
    }
  { \jtermeq
      {{{\Gamma}{{A}{P}}}{Q}}
      {R}
      {\sev{\ctxext{A}{P}}{R}{h}}
      {\sev{P}{R}{\sev{A}{\mprd{P}{R}}{h}}}
    }
\end{align*}

@@

\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\mprd{Q}{R}}}
      {\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\mprd{Q}{R}}}}
      {\ctxwk{A}{\mprd{Q}{S}}}
      {\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{S}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\mprd{Q}{R}}}}
      {\ctxwk{A}{\mprd{Q}{S}}}
      {\ctxwk{A}{\slam{Q}{S}{k}}}
      {\slam{\ctxwk{A}{Q}}{\ctxwk{A}{S}}{\ctxwk{A}{k}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{{\Gamma}{B}}{\mprd{Q}{R}}}{\mprd{Q}{S}}{k}
    }
  { \jtermeq
      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{R}}}
      {\ctxwk{A}{S}}
      {\ctxwk{A}{\sev{Q}{S}{k}}}
      {\sev{\ctxwk{A}{Q}}{\ctxwk{A}{S}}{\ctxwk{A}{k}}}
    }
\end{align*}

@@

\begin{align*}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\subst{x}{P}}}
      {\subst{x}{\mprd{Q}{R}}}
      {\mprd{\subst{x}{Q}}{\subst{x}{R}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
      {\subst{x}{\mprd{Q}{S}}}
      {\mprd{\subst{x}{Q}}{\subst{x}{S}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
      {\subst{x}{\mprd{Q}{S}}}
      {\subst{x}{\slam{Q}{S}{k}}}
      {\slam{\subst{x}{Q}}{\subst{x}{S}}{\subst{x}{k}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{{\Gamma}{A}}{P}}{\mprd{Q}{R}}}{\mprd{Q}{S}}{k}
    }
  { \jtermeq
      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{R}}}
      {\subst{x}{S}}
      {\subst{x}{\sev{Q}{S}{k}}}
      {\sev{\subst{x}{Q}}{\subst{x}{S}}{\subst{x}{k}}}
    } 
\end{align*}

\subsubsection{The dependent function constructor is compatible with weakening}
\begin{align*}
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jfam{{{\Gamma}{Q}}{P}}{R}
    }
  { \jfameq
      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}
      {\mprd{A}{\ctxwk{Q}{R}}}
      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{R}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}{\mprd{A}{\ctxwk{Q}{R}}}}
      {\mprd{A}{\ctxwk{Q}{S}}}
      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{S}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jterm{{{{\Gamma}{P}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}{\mprd{A}{\ctxwk{Q}{R}}}}
      {\mprd{A}{\ctxwk{Q}{S}}}
      {\slam{A}{\ctxwk{Q}{S}}{\ctxwk{Q}{k}}}
      {\ctxwk{\mprd{A}{Q}}{\slam{A}{S}{k}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
    \jterm{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{R}}}{\mprd{A}{S}}{k}
    }
  { \jtermeq
      {{{{{\Gamma}{A}}{P}}{Q}}{\ctxwk{Q}{R}}}
      {\ctxwk{Q}{S}}
      {\sev{A}{\ctxwk{Q}{S}}{\ctxwk{\mprd{A}{Q}}{k}}}
      {\ctxwk{Q}{\sev{A}{S}{k}}}
    }
\end{align*}

@@

>>>>

\begin{align*}
& \inference
    { \jfam{{\Gamma}{A}}{P}
      }
    { \jfam{\Gamma}{\mprd{A}{P}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jfameq{{\Gamma}{A}}{P}{P'}
        }
      { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd{A'}{P'}}
        }
  \\
& \inference
    { \jfam{{{\Gamma}{A}}{P}}{Q}
      }
    { \jfam{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
        }
      { \jfameq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\mprd[\famsym]{A'}{Q'}}
        }
  \\
& \inference
    { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
      }
    { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\slam{A}{Q}{g}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
        }
      { \jtermeq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\slam{A}{Q}{g}}{\slam{A'}{Q}{g'}}
        }
\end{align*}

@@

\subsubsection{Dependent products are compatible with the empty context}
The empty context can appear in the domain and in the codomain of the dependent
function type constructor. We have the following inference rules explaining
what happens when the empty context appears in the domain:
\begin{align}
& \inference
    { \jfam{\Gamma}{P}
      }
    { \jfameq{\Gamma}{\mprd{\emptyf}{P}}{P}
      }
  \\
& \inference
    { \jfam{{\Gamma}{P}}{Q}
      }
    { \jfameq{{\Gamma}{P}}{\mprd[\famsym]{\emptyf}{Q}}{Q}
      }
  \\
& \inference
    { \jterm{{\Gamma}{P}}{Q}{g}
      }
    { \jtermeq{{\Gamma}{P}}{Q}{\slam{\emptyf}{Q}{g}}{g}
      }
\end{align}
We have the following infernece rules explaining what happens when the empty
context appears in the codomain:
\begin{align}
& \inference
    { \jfam{\Gamma}{A}
      }
    { \jfameq{\Gamma}{\mprd{A}{\emptyf}}{\emptyf}
      }
  \\
& \inference
    { \jfam{{\Gamma}{A}}{P}
      }
    { \jfameq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{\emptyf}}{\emptyf}
      }
\end{align}
Now we see that we can compare $\mprd{A}{P}$ with $\mprd[\famsym]{A}{P}$ by
seeing $P$ as a family over $\ctxext{{\Gamma}{A}}{\emptyf}$. We will require
the following inference rule to be valid:
\begin{equation}
\inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd[\famsym]{A}{P}}
    }
\end{equation}
This rule will allow us to omit the annotation indicating the action on families,
which we will do from now on. We also note that because the empty family over
$\ctxext{\Gamma}{A}$ is mapped to the empty family over $\Gamma$ by the
dependent function constructor, we obtain the important special cases of
lambda abstraction and evaluation that
\begin{align*}
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jterm{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}
    }
\end{align*}
Thus, we retrieve what is originally meant by lambda abstraction.

@@

\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq
      {\Gamma}
      {\mprd{\ctxext{A}{P}}{Q}}
      {\mprd{A}{\mprd{P}{Q}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}
      {\mprd{\ctxext{A}{P}}{R}}
      {\mprd{A}{\mprd{P}{R}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
    }
  { \jtermeq
      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}
      {\mprd{\ctxext{A}{P}}{R}}
      {\slam{\ctxext{A}{P}}{R}{h}}
      {\slam{A}{\mprd{P}{R}}{\slam{P}{R}{h}}}
    }
\end{align*}

@@

\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\mprd{Q}{R}}}
      {\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\mprd{Q}{R}}}}
      {\ctxwk{A}{\mprd{Q}{S}}}
      {\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{S}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\mprd{Q}{R}}}}
      {\ctxwk{A}{\mprd{Q}{S}}}
      {\ctxwk{A}{\slam{Q}{S}{k}}}
      {\slam{\ctxwk{A}{Q}}{\ctxwk{A}{S}}{\ctxwk{A}{k}}}
    }
\end{align*}

@@

\begin{align*}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\subst{x}{P}}}
      {\subst{x}{\mprd{Q}{R}}}
      {\mprd{\subst{x}{Q}}{\subst{x}{R}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
      {\subst{x}{\mprd{Q}{S}}}
      {\mprd{\subst{x}{Q}}{\subst{x}{S}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
      {\subst{x}{\mprd{Q}{S}}}
      {\subst{x}{\slam{Q}{S}{k}}}
      {\slam{\subst{x}{Q}}{\subst{x}{S}}{\subst{x}{k}}}
    }
\end{align*}

\subsubsection{The dependent function constructor is compatible with weakening}
\begin{align*}
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jfam{{{\Gamma}{Q}}{P}}{R}
    }
  { \jfameq
      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}
      {\mprd{A}{\ctxwk{Q}{R}}}
      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{R}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}{\mprd{A}{\ctxwk{Q}{R}}}}
      {\mprd{A}{\ctxwk{Q}{S}}}
      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{S}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jterm{{{{\Gamma}{P}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}{\mprd{A}{\ctxwk{Q}{R}}}}
      {\mprd{A}{\ctxwk{Q}{S}}}
      {\slam{A}{\ctxwk{Q}{S}}{\ctxwk{Q}{k}}}
      {\ctxwk{\mprd{A}{Q}}{\slam{A}{S}{k}}}
    }
\end{align*}

@@

\subsection{The evaluation term}
Until now we have just introduced the dependent function constructor, but we 
have not treated one of its main characteristic, the evaluation map. Usually
in type theory, evaluation is introduces by a rule like
\begin{equation*}
\inference
    { \jterm{\Gamma}{\mprd{A}{P}}{f}
      }
    { \jterm{{\Gamma}{A}}{P}{\mathsf{ev}(f)}
      }
\end{equation*}
or even a rule like
\begin{equation*}
\inference
    { \jterm{\Gamma}{\mprd{A}{P}}{f}
      \jterm{\Gamma}{A}{x}
      }
    { \jterm{\Gamma}{\subst{x}{P}}{\mathsf{ev}(f,x)}
      }
\end{equation*}
involving a term of $A$. We immediately discard the latter as a viable option.
In the current setting however, there is an even better option than the first. 
We need a better option because given a family 
$\jfam{{{\Gamma}{A}}{P}}{Q}$, we will need to have a family over
$\ctxext{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}$ with fibers
$\subst{{x}{\mathsf{ev}(f)}}{{x}{Q}}$ at $f:\mprd{A}{P}$ and $x:A$. In a 
name-free setting, such a family is impossible to obtain when using evaluation
as introduced above. The solution is to introduce evaluation via the rule
\begin{equation}\label{eq:evtm}
\inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jterm
      {{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}}
      {\ctxwk{\mprd{A}{P}}{P}}
      {\evtm{A}{P}}
    }
\end{equation}
The advantages are as follows:
\begin{enumerate}
\item It matches the categorical notion of locally cartesian closedness, where
the unit and the counit consist of morphisms in the category.
\item We now have the family 
\begin{equation*}
\jfam
  {{{\Gamma}{\mprd{A}{P}}}{\ctxwk{\mprd{A}{P}}{A}}}
  {\subst{\evtm{A}{P}}{\ctxwk{\mprd{A}{P}}{Q}}}
\end{equation*}
Using this family, we hope to find the judgmental equality
\begin{equation*}
\jfameq
  {{\Gamma}{\mprd{A}{P}}}
  {\mprd{A}{Q}}
  {\mprd{\ctxwk{\mprd{A}{P}}{A}}{\subst{\evtm{A}{P}}{\ctxwk{\mprd{A}{P}}{Q}}}}
\end{equation*}
expressing the action on families of the dependent function constructor in terms
of its action on contexts. This is in agreement with the usual approach of
\cite{TheBook}, where no action on families of the dependent function type
constructor is needed. We will also be able to use the family
\begin{equation*}
\jfam
  {{\Gamma}{\mprd{A}{P}}}
  {{\ctxwk{\mprd{A}{P}}{A}}{\subst{\evtm{A}{P}}{\ctxwk{\mprd{A}{P}}{Q}}}}
\end{equation*}
to formulate the compatibility of the dependent function constructor with itself.
\end{enumerate}
Of course, we will also need the following convertibility rule before we 
continue:

\begin{rmk}
It is common practice, for example in \cite{TheBook}, to omit explicit notation
for evaluation. In our setting that would mean that we would denote a term 
$\subst{f}{\evtm{A}{P}}$ of $P$ in context $\ctxext{\Gamma}{A}$ by $f$.
However, since the evaluation term is set to be an actual term of the theory,
it seems to be harmful to the exposition of the theory to copy this practice.
\end{rmk}

\subsubsection{The compatibility rules of the evaluation terms}

\paragraph{Compatibility with the empty context}
The following rule explains that the evaluation term is judgmentally equal to
the identity term when the domain of the dependent function constructor is the
empty family over $\Gamma$.
\begin{align*}
& \inference
    { \jfam{\Gamma}{P}
      }
    { \jtermeq
        {{\Gamma}{P}}
        {\ctxwk{P}{P}}
        {\evtm{\emptyf}{P}}
        {\idtm{P}}
      }
\end{align*}

\paragraph{Currying for the evaluation term}
We have asserted that there is a judgmental equality
\begin{equation*}
\jfameq
      {\Gamma}
      {\mprd{\ctxext{A}{P}}{Q}}
      {\mprd{A}{\mprd{P}{Q}}}
\end{equation*}
for every family $Q$ over $\ctxext{{\Gamma}{A}}{P}$. It follows that we have
the judgmental equalities
\begin{align*}
& \ctxwk{\mprd{A}{\mprd{P}{Q}}}{Q}
  \\
& \jdeq
  \subst
    { \evtm{A}{\mprd{P}{Q}}}
    { \ctxwk
        {{\mprd{A}{\mprd{P}{Q}}}{\mprd{P}{Q}}}
        {{\mprd{A}{\mprd{P}{Q}}}{Q}}
      }
  \\
& \jdeq
  \subst
    { \evtm{A}{\mprd{P}{Q}}}
    { \ctxwk
        {\mprd{A}{\mprd{P}{Q}}}
        {{\mprd{P}{Q}}{Q}}
      }
\end{align*}
Since $\evtm{P}{Q}$ is a term of $\ctxwk{\mprd{P}{Q}}{Q}$ in 
context $\ctxext{{{\Gamma}{A}}{\mprd{P}{Q}}}{\ctxwk{\mprd{P}{Q}}{P}}$ we may
require the following compatibility rule:
\begin{align}
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \begin{array}{l}
    \ctxext
      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}{\ctxwk{\mprd{\ctxext{A}{P}}{Q}}{\ctxext{A}{P}}}
      \\
    \jtermeq
      {\qquad}
      {\ctxwk{\mprd{\ctxext{A}{P}}{Q}}{Q}}
      {\evtm{\ctxext{A}{P}}{Q}}
      {\subst{\evtm{A}{\mprd{P}{Q}}}{\ctxwk{\mprd{A}{\mprd{P}{Q}}}{\evtm{P}{Q}}}}
    \end{array}
    }
\end{align}
To get a feel for what it says we substitute terms in both expressions involved
in the following lemma.

\begin{lem}
Consider $f:\mprd{\ctxext{A}{P}}{Q}$, $x:A$ and $u:\subst{x}{P}$ be terms in
context $\Gamma$. Then we have the judgmental equality
\begin{equation*}
\jtermeq
  {\Gamma}
  {\subst{u}{{x}{Q}}}
  {\subst{\tmext{x}{u}}{{f}{\evtm{\ctxext{A}{P}}{Q}}}}
  {\subst{u}{{{x}{{f}{\evtm{A}{\mprd{P}{Q}}}}}{\evtm{\subst{x}{P}}{\subst{x}{Q}}}}}
\end{equation*}
is valid.
\end{lem}

\paragraph{Weakening is compatible with the evaluation term}
\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \begin{array}{l}
    \ctxext
        { { { {\Gamma}
              {A}
              }
            { \ctxwk{A}{B}
              }
            }
          { \mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}
            }
          }
        { \ctxwk{\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}{\ctxwk{A}{Q}}
          }
      \\
    \jtermeq
      {\qquad}
      {\ctxwk{\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}{\ctxwk{A}{R}}}
      {\evtm{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
      {\ctxwk{A}{\evtm{Q}{R}}}  
    \end{array}
    }
\end{align*}

\paragraph{Substitution is compatible with the evaluation term}
\begin{align*}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \begin{array}{l}
    \ctxext
      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
      {\ctxwk{\mprd{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{Q}}}
      \\
    \jtermeq
      {\qquad}
      {\ctxwk{\mprd{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{R}}}
      {\subst{x}{\evtm{Q}{R}}}
      {\evtm{\subst{x}{Q}}{\subst{x}{R}}}
    \end{array}
    }  
\end{align*}

\paragraph{The evaluation term is compatible with weakening}
\begin{align*}
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
    \jterm{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{R}}}{\mprd{A}{S}}{k}
    }
  { \jtermeq
      {{{{{\Gamma}{A}}{P}}{Q}}{\ctxwk{Q}{R}}}
      {\ctxwk{Q}{S}}
      {\sev{A}{\ctxwk{Q}{S}}{\ctxwk{\mprd{A}{Q}}{k}}}
      {\ctxwk{Q}{\sev{A}{S}{k}}}
    }
\end{align*}

\subsection{Composition of dependent functions}
